### 2024-02-27 manuscript.Rmd
27/2

---
title: "Opportunities for improvement in adult trauma patients admitted to the intensive care unit: A registry-based study"
output: html_document
author: "Elin Sun Cao"
bibliography: bibliography.bib
csl: vancouver.csl
editor_options: 
  markdown: 
    wrap: 72
---

# Abstract

## Background

Trauma is a leading cause of mortality and morbidity worldwide. A
cornerstone in developing the trauma quality improvement programs is
multidisciplinary mortality and morbidity review. The purpose of these
is to identify opportunities for improvement (OFI), which are necessary
in order to implement corrective actions. Many patients with severe
trauma are admitted to the intensive care unit (ICU), but little is
known about opportunities for improvement in this group of patients. The
aim of this study is to characterise OFI in adult trauma patients
admitted to the ICU and assess how patient factors are associated with
OFI in these patients.

## Methods

This was a registry-based study using all patients between 2014 and 2022
from the Karolinska University Hospital in Solna who had been admitted
to the ICU and who were reviewed regarding the presence of OFI by the
mortality and morbidity conference. Bi-variable and multivariable
logistic regression was used to assess the following patient and process
factors and OFI among this group of patients: gender, age, if the
patient was intubated or not and how many days, respiratory rate,
systolic blood pressure, Glasgow Coma Scale (GCS), Injury Severity Score
(ISS), on call, time from arrival at the hospital until first computed
tomography (CT), ASA preinjury and survival after 30 days.

## Results

## Conclusion

# Abbreviations

CT = computed tomography

GCS = Glasgow Coma Scale

GOS = Glasgow Outcome Score

ISS = Injury Severity Score

NISS = New Injury Severity Score

OFI = Opportunity For Improvement

RTS = Revised Trauma Score

TRISS = Trauma and Injury Severity Score

# Introduction

Trauma --- the clinical entity of injury and the body's associated
response --- is a leading cause of mortality and morbidity worldwide.
According to WHO, injury-related deaths constitute 4% of all deaths and
take the lives of 4,4 million people yearly worldwide - more than
HIV/AIDS, malaria, maternal mortality and tuberculosis combined. The top
3 of 5 causes of death in people aged 5-29 are injury-related; road
traffic injuries, homicide and suicide. Each year, tens of millions of
people suffer non-fatal injuries leading to treatment at hospitals in
emergency departments and acute care visits. These injuries often lead
to temporary or permanent disability and the need for medical and mental
rehabilitation. Trauma exposure at a young age increases the risk for
mental illness and suicide leading to increased risk of an unhealthy
lifestyle associated with smoking, alcohol and substance abuse, chronic
disease and cancer as well as societal problems such as crime, poverty
and violence. The burden of disease from injuries vary depending on
certain conditions; socioeconomic factors, gender, age and country. Men
suffer from injuries more often than women. Across all age groups
worldwide, the top three leading causes of injury among males are road
traffic injuries, homicide and suicide; whereas the top three leading
causes of injury among women are road traffic injuries, falls and
suicide [@1,@2].

According to the annual report from the Swedish Trauma registry
(SweTrau) 2022, trauma was more common among men than women in Sweden
--- 65,3% men compared to 34,7% women. The most affected age group was
the working population. The leading causes of trauma included; road
traffic injuries, falls, weapons and stabs and blows. Previous years,
road traffic injuries have been the leading cause of injury. However,
since 2022 road traffic injuries and falls account for about the same
proportion of mechanisms of injury. There are several variables that can
be studied to assess primary patient care, for example time to first CT
and acute interventions. Median time to first CT at the university
hospitals of Sweden was 30 minutes in unconscious patients with GCS \<
9. Acute interventions is defined as interventions performed immediately
or within 24 hours after arrival to a hospital. The most common acute
interventions in Swedish hospitals were: thoracic drains, wound
revisions and fracture surgery followed by laparotomy with haemostasis,
radiological interventions such as embolisation and stents, and
thoracotomy [@3].

Albaaj et al. recently analyzed opportunities for improvement (OFI) in
relation to different patient and process factors based on data from the
Swedish trauma registry and data from the Karolinska University Hospital
in Solna. The study indicated that patients with moderate to severe
trauma are at the highest odds of OFI [@4]. According to SweTrau 2022
41,1% of patients with New Injury Severity Score (NISS) \> 15 who were
at a university hospital were admitted to the ICU. The proportion of
severely injured admitted to the ICU has decreased in recent years,
however; patients with NISS \> 15 still constitute a large proportion of
the trauma patients treated in the ICU. A way to measure functional
level after hospital discharge is Glasgow Outcome Scale (GOS). According
to GOS, among patients with NISS \> 15 17% were deceased, 20% suffered
severe disability, 49% suffered moderate disability and 14% had good
recovery in the university hospitals of Sweden [@3].

For all injuries, quality hospital care can reduce the amount of short-
and long-term disability. Therefore, improving planning, access and
organisation of trauma care systems including pre-hospital care,
interhospital transportation, telecommunication and in-hospital care all
play a central role in reducing overall impact after trauma [@1].
Disability-adjusted life-years (DALYs) is a way to measure
post-discharge quality of life. Trauma accounts for one-tenth of DALYs,
most of the burden in low- and middle-income countries (LMIC) [@5, @6].
Trauma costs approximately 3% of gross domestic product (GDP) yearly
globally, as it mostly affects the working population [@7].

## Trauma quality improvement programs

Many initiatives for quality improvement programs have been seen in
trauma and critical care based on the paradigm of Donabedian [@8],
stating that implementing proven structures and processes of care is the
most effective way to improve outcomes. One example of a structure and
process intervention in the intensive care unit is highly-staffed ICUs
that has been associated with improved outcomes, as well as bundles to
recognize and treat common conditions such as those for treating sepsis
and ventilator-associated pneumonia [@9]. The intention of quality
improvement is objective improvement in regard to preventability. A
cornerstone in developing the trauma quality improvement programs is
therefore multidisciplinary mortality and morbidity review. The purpose
of these is to discuss preventable deaths and identifying OFIs in trauma
care regarding structure as well as clinical processes. This way,
identifying OFIs will provide guidance as to what to focus specific
efforts on. There are several other methods to improve trauma care.
However, mortality and morbidity review has been shown to be a key
method of addressing all necessary components of trauma improvement
[@10].

Trauma audit filters are descriptions of specific actions that should be
taken, time frames in which tests or treatments should be provided, or
outcomes that are expected to occur in injured patients. Assessing
patients whose care falls outside the frames of these audit filters and
providing feedback to the clinicians involved in the patient's care
provides means of correcting errors and improving future trauma care.
The audit filters were introduced by the American College of Surgeons
1995; however, they have been criticized as they may not be directly
linked to meaningful clinical or patient-centered outcomes [@11]. At the
Karolinska University Hospital in Solna local audit filters are used.
These local audit filters include: systolic blood pressure under 90, GCS
\< 9 and not intubated, ISS \> 15 but not admitted to the ICU, ISS \> 15
and no trauma team activation, no anticoagulation treatment within 72 h
after traumatic brain injury, time to acute intervention more than 60
minutes, time to CT more than 30 minutes, liver or spleen injury,
cardiopulmonary resuscitation with thoracotomy, massive transfusion and
death within 30 days after trauma.

## Injury score systems

When a patient first arrives at the hospital, injury score systems are
used to assess how urgent the need of interventions to a patient is.
There are several injury score systems: Injury Severity Score (ISS), New
Injury Severity Score (NISS), Revised Trauma Score (RTS) and Trauma and
Injury Severity Score (TRISS).

ISS is an anatomical scoring system providing an overall score for
patients with multiple injuries. Each injury is assigned an abbreviated
injury scale (AIS) score and is allocated to one of six body regions.
The three most severely injured body regions have their scores squared
and added together to produce an ISS score between 0 and 75. If an
injury is assigned an AIS of 6 the ISS score is automatically assigned
to 75. Major trauma is considered when ISS \> 15. The major drawback of
ISS is that it only considers one injury for each body region.
Therefore, if several severe injuries are in the same body region, they
are not accounted for. NISS counters the problem of ISS being the sum of
the squares of the three most severe injuries, regardless of body region
injured. NISS is therefore equal to or higher than ISS.

RTS is a physiological scoring system scored from the data obtained in
the first set of data obtained on the patient and consists of Glasgow
Coma Scale (GCS), respiratory rate and systolic blood pressure. RTS is
heavily weighted towards GCS compared to respiratory rate and systolic
blood pressure. It has been demonstrated to show accuracy at predicting
death.

TRISS has a maximum prediction in outcome when compared with the other
scores. The score determines the probability of survival based on a
formula consisting of ISS, RTS and the patient's age. TRISS is therefore
both an anatomical and physiological scoring system [@12].

## Preventable deaths in patients admitted to the ICU

There are many studies on preventable deaths in trauma patients treated
in the ICU both regarding the medical aspect as well as the human
aspect. According to a study from Japan on the medical aspect, the
primary medical causes of death after trauma are haemorrhage,
sepsis/multiorgan failure and central nervous system injury. Delayed
hemostatic procedures and transfusions were common areas for improvement
in the hospital stage [@13]. The most prominent OFI in regard to
haemorrhage was decision-making compared to errors in technical skill.
The OFIs frequently involved the decision between radiology, surgery and
further investigation [@14].

A study from The University of Pennsylvania looked at the human aspect
of OFIs in the ICU. The ICU of this hospital uses telemedicine and
telemonitoring to review critical patient events. The areas of
improvement they identified were: team dynamics and communication, use
of best practices, increased and standardized access to emergency
resources and improvement of procedural technique. They recommend other
ICUs with telemedicine capacity to adopt a similar strategy in order to
present OFIs and educational discussions [@15]. Another study looked at
how critical information was handled in the ICU and found that
laboratory values and test results were the most frequently lost items.
This points at the importance of communication and handling of critical
information at handoffs of care [@16].

Preventability in studies is often defined as an event that would not
have occurred had the patient received ordinary standards of care.
Evaluating preventable deaths in patients admitted to the ICU can be a
challenge as the patients have extensive medical history and it is
harder to understand what the standard of care is for those patients.
Furthermore, even though these patients receive the best of care, these
patients still have a higher risk of developing complications or dying.
Some might even be in such a bad condition that they die just when they
have been admitted to the ICU from the ward or emergency care [@9].

Many patients with severe trauma are admitted to the intensive care unit
(ICU), but little is known about opportunities for improvement in this
group of patients. This being due to the patients admitted to the ICU
being more complex. Identifying OFIs in this group can therefore be more
of a challenge. There are existing studies on different aspects of
trauma care improvement, but studies focusing mainly on OFI in the ICU
are few.

## Aim

The aim of this study is to characterise OFI in adult trauma patients
admitted to the ICU and assess how patient factors are associated with
OFI in these patients.

# Methods

## Study design

We conducted a retrospective single-centre cohort study based on data
from the Karolinska University Hospital in Solna and from the Swedish
Trauma registry. We used R for all statistical analyses [@19]. For
patients where hospital values for systolic blood pressure, respiratory
rate or GCS score were missing, prehospital values were used instead. A
complete case analysis was conducted after handling missing values. We
present sample characteristics using descriptive statistics.

## Setting

The Karolinska University Hospital in Solna is a level 1 trauma centre
based on the definition by the American College of Surgeons (ACS)
Committee on Trauma [@17], which means that all patients in Stockholm
with severe or suspected severe trauma are sent there. The hospital has
direct access to radiology, surgery, emergency medicine, intensive care
and consultants from relevant specialities [@18]. All patients are
included in a mortality and morbidity screening combining individual
review and specific audit filters by specialized nurses. The patients
who are identified as having higher probability for OFI are discussed in
a multidisciplinary conference held every 6-8 weeks. Examples of OFI
include the need for better organization or more senior members taking
part in the trauma care. The presence or absence of OFI is decided among
the participants of the conference and then recorded in the trauma care
quality database.

## Participants

The trauma registry includes all patients admitted with trauma team
activation, regardless of Injury Severity Score (ISS), as well as
patients admitted without trauma team activation but found to have an
ISS of more than 9. We included all patients who had been included in
the morbidity and mortality screening process between January 1, 2014
and February 1, 2023 and who were also admitted to the ICU. We excluded
patients who were younger than 15 years and patients who were dead on
arrival.

## Variables and data sources/measurements

### Study outcome

The study outcome will be on the presence of OFI in adult patients
treated in the ICU based on patient and process factors.

### Patient and process factors

The quantitative variables were factors chosen from the trauma registry,
based on the locally used audit filters, standard epidemiological
factors and factors registered in the Swedish Trauma Registry. The
categorical factors were sex, age, if the patient was intubated or not
and how many days, respiratory rate, systolic blood pressure, Glasgow
Coma Scale (GCS), Injury Severity Score (ISS), on call, time from
arrival at the hospital until first computed tomography (CT), ASA
preinjury and survival after 30 days. On duty was defined as before 8.00
a.m., after 5 p.m., or during a weekend - defined as Saturday or Sunday.

## Study size

There were ... patients in the trauma registry between January 1 2014
and February 1 2023. ... with trauma team activation regardless of ISS
score, ... patients admitted without trauma team activation but with ISS
score \> 9. .... patients were included in the morbidity and mortality
screening process between 2014 and 2022. ... of these patients were also
admitted to the ICU. ... patients were excluded as they were younger
than 15 years and ... patients were excluded because they were dead on
arrival. This leaves us with 167 outcomes.

## Statistical methods

Bivariable logistic regression was used to determine unadjusted
associations and multivariable logistic regression to determine adjusted
associations between patient and process factors and OFI. Odds ratios
(OR) with associated 95% confidence intervals was used with a 5%
significance level.

## Ethical considerations

There are several ethical aspects to take into consideration regarding
this study: voluntary participation, informed consent, anonymity,
confidentiality, potential for harm and results communication.

As described previously, all patients with trauma team activation or
high ISS scores are included in the database used for this study. Having
suffered severe trauma might not be something the patients or their
relatives would want to be public information as the collected data
contain for instance vital parameters and demise as well as type of
trauma and time of arrival to the hospital, which might be sensitive
information. Therefore, the ethical aspect of voluntary participation
and informed consent has to be considered. The data in national
databases is collected from all patients registered and is not based on
consent, even though it might be informed, which is a drawback.

To respect the anonymity and confidentiality aspect the statistical
analysis was first conducted on synthetic data and later implemented on
the data collected from the trauma registry and the trauma care quality
database in order to protect personal numbers and their associated
information. The anonymity might ease the discomfort among the involved
patients and their relatives.

There is no direct potential for harm by being included in the study as
no intervention is made and it is an observational study.

In regard to results communication, the results from research based on
the data as well as a compilation of the original data is public. The
data being accessible for research does not directly benefit the
affected individual; however, it contributes greatly to the development
of the hospital care in Sweden, making it more efficient and equal.
Studies performed on data from the Karolinska University Hospital in
Solna are made to improve local trauma care systems. This way, it can
also benefit the society as improved local trauma care decreases trauma
death rate and improves quality of care and thus patient safety.

Based on these arguments it can be concluded that the advantages
outweigh the disadvantages. The advantages being the contribution to the
development of trauma systems which long-term improves patient safety
and equality, and that the study protects the individual's anonymity and
confidentiality by using synthetic data and later implementing it on the
real data. The disadvantage being the potential emotional harm from the
absence of voluntary participation.

The study was approved by the Swedish Ethical Review Authority, approval
numbers 2023-02975-02, 2021-03531, 2021-02541.

# Results

14,022 patients were included in the trauma registry and trauma care
quality database between 2017 and 2023. 12,278 patients were excluded,
leaving a total of 1,744 patients eligible for the study. 11,343 were
excluded because they were not treated in the ICU. 3 patients were
excluded because they were under 15 years old. 6 patients were excluded
because they were dead on arrival, and 926 patients were excluded
because there was no data on the presence or absence OFI.

[n] patients were excluded because there was missing data, leaving a
total of [n] patients included in the study.

Table 1 present sample characteristics. [n{%}] patients were male and
[n{%}] patients were female. [n{%}] patients had no intubation. Most
patients of the intubated ones were intubated for xx days [n{%}]. The
number of patients who arrived outside working hours - on Saturday or
Sunday or after 5 pm or before 8 am were [n{%}]. [n{%}] patients had an
ASA-score of 1-2 before arrival to the hospital, and [n{%}] patients had
an ASA-score of 3-7 before arrival to the hospital. [n{%}] patients died
within 30 days after hospitalization.

Table 2 shows the unadjusted and adjusted associations of the selected
patient and process factors in relation to OFI. In the unadjusted
analysis the factors that were significantly associated with the
presence of OFI were: ... Male
`r inline_text(table3a, variable = Sex, level = "Male")` Female
`r inline_text(table3a, variable = Sex, level = "Female")` Age
`r inline_text(table3a, variable = Age)` No intubation
`r inline_text(table3a, variable = Intubation, level = "No intubation")`
Respiratory rate `r inline_text(table3a, variable = RespiratoryRate)`
Systolic blood pressure
`r inline_text(table3a, variable = SystolicBloodPressure)` Glasgow Coma
Scale `r inline_text(table3a, variable = GlasgowComaScale)` Jour
`r inline_text(table3a, variable = Jour, level = "Yes")` Time to first
CT `r inline_text(table3a, variable = TimeFCT)` Time to first
intervention `r inline_text(table3a, variable = TimeFInt)` ASA preinjury
1-2 `r inline_text(table3a, variable = ASApreinjury, level = "ASA 1-2")`
Death within 30 days
`r inline_text(table3a, variable = Survival, level = "Dead")`

In the adjusted analysis the following factors demonstrated significant
association with OFI: ...

x was significantly associated with increased odds of OFI in both the
unadjusted and adjusted analyses.

x were not significantly associated with OFI in the unadjusted or the
adjusted analyses.

## Participants

<!-- Flowchart -->

## Descriptive data

<!-- Sample characteristics - Table1, do I need to display number of missing for each variable? -->

## Main results

<!--  Adjusted and unadjusted analysis - Table 2 -->

# Discussion

### Key results

We found that --- was associated with higher odds of OFI. Previous
research ...

These findings could be because ...

\-- was associated with OFI in the unadjusted analysis but not in the
adjusted analysis.

1.  add sources from previous studies

2.  studies to back up the results from the analyses

### Limitations

There were several limitations in our study. The selection for review
for the morbidity and mortality review relied mostly on the local audit
filters used at the Karolinska Institute Hospital in Solna. This leads
to a risk of misplacement where some patients with OFI were not selected
by for the review based on the audit filters although there was an OFI.

Another limtation is that some relevant factors, such as ----, were not
part of the Swedish Trauma registry and therefore not adjusted for in
the analysis.

The study was a single-center study and illustrate the situation in
Stockholm. More study is needed on the area of OFI in patients admitted
to the ICU. A suggestion would be to conduct a multi-center study on
this group of patients to further investigate areas for improvement and
to further develop trauma quality programs.

Future studies could also focus on ---- which was seen to be associated
with OFI in these patients.

<!--Lägg till 2 st förslag på explicita fortsättningsstudier! -->

### Generalisability

The study is conducted on --- outcomes. There were few outcomes because
we included only those who were admitted to the ICU, \> 14 years old,
not dead on arrival and had a registration for the presence or absence
of OFI. The external validity is therefore compromised due to the
limited size of the Swedish database. However, ...

# Conclusion

---- among patients admitted to the ICU have higher odds of OFI. These
factors can be seen as new targets for further reseach to further
supplement trauma quality improvement programs.

<!-- Give a short conclusion, corresponding to the aim -->

# References

<!-- Do not edit by hand, references will be inserted and formatted automatically once you knit this document -->