### 2024-03-25 manuscript.Rmd
25/3

---
title: "Opportunities for improvement in adult trauma patients admitted to the intensive care unit: A registry-based study"
output: "html_document"
author: "Elin Sun Cao"
bibliography: bibliography.bib
csl: vancouver.csl
editor_options: 
  markdown: 
    wrap: 72
---

# Abstract

## Bakgrund

Trauma är en ledande orsak till mortalitet och morbiditet i världen. En
hörnsten i utvecklingen av förbättringsprogram för trauma är
multidisciplinära mortalitets- och morbiditetskonferenser. Syftet med
dessa är att identifiera opportunities for improvement (OFI) som är en
viktig del i det riktade förbättringsarbetet av traumahandläggning.
Många patienter med svåra trauman blir inlagda på
intensivvårdsavdelningen (IVA). Trots detta är kännedomen om
förbättringsmöjligheter hos denna patientgrupp inte heltäckande i
nuläget. Syftet med denna studie är att karaktärisera OFI hos vuxna
traumapatienter som varit inlagda på IVA och bedöma hur patient- och
processfaktorer är associerade med OFI hos denna patientgrupp.

## Metod

Detta är en registerbaserad studie som inkluderar traumapatienter från
2014 till 2023 från Karolinska Universitetssjukhuset i Solna som har
varit inlagda på IVA och som har granskats avseende förekomsten av OFI
vid en mortalitets- och morbiditetskonferens. Bivariabel och
multivariabel logistisk regression användes för att bedöma sambandet
mellan följande patient- och processfaktorer samt OFI bland denna
patientgrupp: kön, ålder, om patienten var intuberad eller ej och hur
många dagar, Revised Trauma Score (RTS), Injury Severity Score (ISS),
jourtid vid ankomst till sjukhuset, tid från ankomst till sjukhuset till
första datortomografi (DT), dagar på IVA, ASA-värde före trauma och
överlevnad efter 30 dagar.

## Resultat

OFI identifierades hos 143 av 1449 patienter. Högre Revised Trauma Score
(RTS) var associerat med högre odds för OFI, ≤ 7 dagar på
intensivvårdsavdelningen (IVA) var associerat med lägre odds för OFI och
död inom 30 dagar efter sjukhusvistelse var associerat med högre odds
för OFI med statistisk signifikans i den ojusterade analysen. Högre
Revised Trauma Score (RTS) var associerat med högre odds för OFI, och ≤
7 dagar på IVA var associerat med lägre odds för OFI med statistisk
signifikans i den justerade analysen.

## Slutsats

De patient- och processfaktorer som var signifikant associerade med OFI
visade att högre RTS var associerat med högre odds för OFI och att ≤ 7
dagar på IVA var associerat med lägre odds för OFI.

## Background

Trauma is a leading cause of mortality and morbidity worldwide. A
cornerstone in developing the trauma quality improvement programs is
multidisciplinary mortality and morbidity review. The purpose of these
is to identify opportunities for improvement (OFI), which are necessary
in order to implement corrective actions. Many patients with severe
trauma are admitted to the intensive care unit (ICU), but little is
known about opportunities for improvement in this group of patients. The
aim of this study is to characterise OFI in adult trauma patients
admitted to the ICU and assess how patient factors are associated with
OFI in these patients.

## Methods

This was a registry-based study using all patients between 2014 and 2023
from the Karolinska University Hospital in Solna who had been admitted
to the ICU and who were reviewed regarding the presence of OFI by the
mortality and morbidity conference. Bivariable and multivariable
logistic regression was used to assess the following patient and process
factors and OFI among this group of patients: gender, age, if the
patient was intubated or not and how many days, Revised Trauma Score
(RTS), Injury Severity Score (ISS), arrival outside of working hours,
time from arrival at the hospital until first computed tomography (CT),
days in the ICU, ASA preinjury and survival after 30 days.

## Results

OFI was identified in 143 of 1449 patients. Higher Revised Trauma Score
(RTS) was associated with higher odds of OFI, ≤ 7 days in the intensive
care unit (ICU) was associated with lower odds of OFI, and death within
30 days after hospitalization was associated with higher odds of OFI
with statistical significance in the unadjusted analysis. Higher Revised
Trauma Score (RTS) was associated with higher odds of OFI, and ≤ 7 days
in the ICU was associated with lower odds of OFI with statistical
significance in the adjusted analysis.

## Conclusion

The patient and process factors associated with OFI indicated that
higher RTS was associated with higher odds of OFI, and ≤ 7 days in the
ICU was associated with lower odds of OFI.

# Abbreviations

APACHE = Acute Physiology and Chronic Health Evaluation

ASA score = American Society of Anesthesiologists score

CT = computed tomography

GCS = Glasgow Coma Scale

GOS = Glasgow Outcome Score

ISS = Injury Severity Score

NISS = New Injury Severity Score

OFI = Opportunity For Improvement

RR = respiratory rate

RTS = Revised Trauma Score

SBP = systolic blood pressure

TRISS = Trauma and Injury Severity Score

# Introduction

Trauma --- the clinical entity of injury and the body's associated
response --- is a leading cause of mortality and morbidity worldwide.
According to WHO, injury-related deaths constitute 4% of all deaths and
take the lives of 4,4 million people yearly worldwide - more than
HIV/AIDS, malaria, maternal mortality and tuberculosis combined. The top
3 of 5 causes of death in people aged 5-29 are injury-related; road
traffic injuries, homicide and suicide. Each year, tens of millions of
people suffer non-fatal injuries leading to treatment at hospitals in
emergency departments and acute care visits. These injuries often lead
to temporary or permanent disability and the need for medical and mental
rehabilitation. Trauma exposure at a young age increases the risk for
mental illness and suicide leading to increased risk of an unhealthy
lifestyle associated with smoking, alcohol and substance abuse, chronic
disease and cancer as well as societal problems such as crime, poverty
and violence. The burden of disease from injuries vary depending on
certain conditions; socioeconomic factors, gender, age and country. Men
suffer from injuries more often than women. Across all age groups
worldwide, the top three leading causes of injury among males are road
traffic injuries, homicide and suicide; whereas the top three leading
causes of injury among women are road traffic injuries, falls and
suicide [@1,@2].

According to the annual report from the Swedish Trauma registry
(SweTrau) 2022, trauma was more common among men than women in Sweden
--- 65,3% men compared to 34,7% women. The most affected age group was
the working population. The leading causes of trauma included; road
traffic injuries, falls, weapons and stabs and blows. Previous years,
road traffic injuries have been the leading cause of injury. However,
since 2022 road traffic injuries and falls account for about the same
proportion of mechanisms of injury. There are several variables that can
be studied to assess primary patient care, for example time to first CT
and acute interventions. Median time to first CT at the university
hospitals of Sweden was 30 minutes in unconscious patients with GCS \<
9. Acute interventions is defined as interventions performed immediately
or within 24 hours after arrival to a hospital. The most common acute
interventions in Swedish hospitals were: thoracic drains, wound
revisions and fracture surgery followed by laparotomy with haemostasis,
radiological interventions such as embolisation and stents, and
thoracotomy [@3].

Albaaj et al. recently analyzed opportunities for improvement (OFI) in
relation to different patient and process factors based on a selection
of data from SweTrau including the patients admitted to the Karolinska
University Hospital in Solna following trauma. The study indicated that
patients with moderate to severe trauma are at the highest odds of OFI
[@4]. According to SweTrau 2022 41,1% of patients with New Injury
Severity Score (NISS) \> 15 who were at a university hospital were
admitted to the ICU. The proportion of severely injured admitted to the
ICU has decreased in recent years, however; patients with NISS \> 15
still constitute a large proportion of the trauma patients treated in
the ICU. A way to measure functional level after hospital discharge is
Glasgow Outcome Scale (GOS). According to GOS, among patients with NISS
\> 15 17% were deceased, 20% suffered severe disability, 49% suffered
moderate disability and 14% had good recovery in the university
hospitals of Sweden [@3].

For all injuries, quality hospital care can reduce the amount of short-
and long-term disability. Therefore, improving planning, access and
organisation of trauma care systems including pre-hospital care,
interhospital transportation, telecommunication and in-hospital care all
play a central role in reducing overall impact after trauma [@1].
Disability-adjusted life-years (DALYs) is a way to measure
post-discharge quality of life. Trauma accounts for one-tenth of DALYs,
most of the burden in low- and middle-income countries (LMIC) [@5, @6].
Trauma costs approximately 3% of gross domestic product (GDP) yearly
globally, as it mostly affects the working population [@7].

## Trauma quality improvement programs

Many initiatives for quality improvement programs have been seen in
trauma and critical care based on the paradigm of Donabedian [@8],
stating that implementing proven structures and processes of care is the
most effective way to improve outcomes. One example of a structure and
process intervention in the intensive care unit is highly-staffed ICUs
that has been associated with improved outcomes, as well as bundles to
recognize and treat common conditions such as those for treating sepsis
and ventilator-associated pneumonia [@9]. The intention of quality
improvement is objective improvement in regard to preventability. A
cornerstone in developing the trauma quality improvement programs is
therefore multidisciplinary mortality and morbidity review. The purpose
of these is to discuss preventable deaths and identifying OFIs in trauma
care regarding structure as well as clinical processes. This way,
identifying OFIs will provide guidance as to what to focus specific
efforts on. There are several other methods to improve trauma care.
However, mortality and morbidity review has been shown to be a key
method of addressing all necessary components of trauma improvement
[@10].

Trauma audit filters are descriptions of specific actions that should be
taken, time frames in which tests or treatments should be provided, or
outcomes that are expected to occur in injured patients. Assessing
patients whose care falls outside the frames of these audit filters and
providing feedback to the clinicians involved in the patient's care
provides means of correcting errors and improving future trauma care.
The audit filters were introduced by the American College of Surgeons
1995; however, they have been criticized as they may not be directly
linked to meaningful clinical or patient-centered outcomes [@11]. At the
Karolinska University Hospital in Solna local audit filters are used.
These local audit filters include: systolic blood pressure under 90, GCS
\< 9 and not intubated, ISS \> 15 but not admitted to the ICU, ISS \> 15
and no trauma team activation, no anticoagulation treatment within 72 h
after traumatic brain injury, time to acute intervention more than 60
minutes, time to CT more than 30 minutes, liver or spleen injury,
cardiopulmonary resuscitation with thoracotomy, massive transfusion and
death within 30 days after trauma.

## Injury score systems

When a patient first arrives at the hospital, injury score systems are
used to assess how urgent the need of interventions to a patient is.
There are several injury score systems: Injury Severity Score (ISS), New
Injury Severity Score (NISS), Revised Trauma Score (RTS) and Trauma and
Injury Severity Score (TRISS).

ISS is an anatomical scoring system providing an overall score for
patients with multiple injuries. Each injury is assigned an abbreviated
injury scale (AIS) score and is allocated to one of six body regions.
The three most severely injured body regions have their scores squared
and added together to produce an ISS score between 0 and 75. If an
injury is assigned an AIS of 6 the ISS score is automatically assigned
to 75. Major trauma is considered when ISS \> 15. The major drawback of
ISS is that it only considers one injury for each body region.
Therefore, if several severe injuries are in the same body region, they
are not accounted for. NISS counters the problem of ISS being the sum of
the squares of the three most severe injuries, regardless of body region
injured. NISS is therefore equal to or higher than ISS.

RTS is a physiological scoring system scored from the data obtained in
the first set of data obtained on the patient and consists of Glasgow
Coma Scale (GCS), respiratory rate and systolic blood pressure. RTS is
heavily weighted towards GCS compared to respiratory rate and systolic
blood pressure. It has been demonstrated to show accuracy at predicting
death. Each of the mentioned variables get a corresponding code value
between 0 and 4 based on the value of GCS, RR and SBP. The formula for
RTS is RTS = (0.9368 x GCS code value) + (0.7326 x SBP code
value) + (0.2908 x RR code value) [@20].

TRISS has a maximum prediction in outcome when compared with the other
scores. The score determines the probability of survival based on a
formula consisting of ISS, RTS and the patient's age. TRISS is therefore
both an anatomical and physiological scoring system [@12].

## Classifications systems for patient's general health

Another way to assess the state of a patient in addition to the injury
score systems are classification systems for patient's general health.
Two of these are ASA score and APACHE II.

American Society of Anesthesiologists (ASA) classification was
originally created to establish a scoring system for evaluation of a
patient's general health and comorbidities before an operative procedure
[@39]. However, recent studies have also showed ASA class to be a
significant predictor for length of stay, indicating that ASA score can
be used as predictor in other aspects of hospital care as well [@40].

Acute Physiology and Chronic Health Evaluation (APACHE) II is the most
commonly used severity-of-disease scoring system in ICUs all around the
world. The worst value for each physiological variabel is converted to
an integer score from 0 to 71 within the first 24 hours of admittance.
Higher score indicates more severe disease and higher risk for
mortality. The score has not been validated for use on patient's under
16 years of age. APACHE II has shown to be a good score for
discriminating between survivors and non-survivors [@41].

## Preventable deaths in patients admitted to the ICU

There are many studies on preventable deaths in trauma patients treated
in the ICU both regarding the medical aspect as well as the human
aspect. According to a study from Japan on the medical aspect, the
primary medical causes of death after trauma are haemorrhage,
sepsis/multiorgan failure and central nervous system injury. Delayed
hemostatic procedures and transfusions were common areas for improvement
in the hospital stage [@13]. The most prominent OFI in regard to
haemorrhage was decision-making compared to errors in technical skill.
The OFIs frequently involved the decision between radiology, surgery and
further investigation [@14].

A study from The University of Pennsylvania looked at the human aspect
of OFIs in the ICU. The ICU of this hospital uses telemedicine and
telemonitoring to review critical patient events. The areas of
improvement they identified were: team dynamics and communication, use
of best practices, increased and standardized access to emergency
resources and improvement of procedural technique. They recommend other
ICUs with telemedicine capacity to adopt a similar strategy in order to
present OFIs and educational discussions [@15]. Another study looked at
how critical information was handled in the ICU and found that
laboratory values and test results were the most frequently lost items.
This points at the importance of communication and handling of critical
information at handoffs of care [@16].

Preventability in studies is often defined as an event that would not
have occurred had the patient received ordinary standards of care.
Evaluating preventable deaths in patients admitted to the ICU can be a
challenge as the patients have extensive medical history and it is
harder to understand what the standard of care is for those patients.
Furthermore, even though these patients receive the best of care, these
patients still have a higher risk of developing complications or dying.
Some might even be in such a bad condition that they die just when they
have been admitted to the ICU from the ward or emergency care [@9].

Many patients with severe trauma are admitted to the intensive care unit
(ICU), but little is known about opportunities for improvement in this
group of patients. This being due to the patients admitted to the ICU
being more complex. Identifying OFIs in this group can therefore be more
of a challenge. There are existing studies on different aspects of
trauma care improvement, but studies focusing mainly on OFI in the ICU
are few.

## Aim

The aim of this study is to characterise OFI in adult trauma patients
admitted to the ICU and assess how patient factors are associated with
OFI in these patients.

# Methods

## Study design

We conducted a retrospective single-centre cohort study based on data
from the Karolinska University Hospital in Solna and from the Swedish
Trauma registry. We used R for all statistical analyses [@19].

For patients where hospital values for systolic blood pressure,
respiratory rate or GCS score were missing from the emergency
department, prehospital values were used instead. If the patient was
intubated we automatically assumed that GCS was equal to 3 and RR was
equal to 0. For the RTS-variable, each of the variables GCS, SBP and RR
were then converted to a RTS-code between 0 and 4 based on what the
values of GCS, SBP and RR were respectively. RTS was then calculated
using the formula for RTS based on the combination of the RTS-codes
assigned to each of the previously listed variables.

A complete case analysis was conducted after handling missing values. We
present sample characteristics using descriptive statistics.

## Setting

The Karolinska University Hospital in Solna is a level 1 trauma centre
based on the definition by the American College of Surgeons (ACS)
Committee on Trauma [@17], which means that all patients in Stockholm
with severe or suspected severe trauma are sent there. The hospital has
direct access to radiology, surgery, emergency medicine, intensive care
and consultants from relevant specialities [@18]. All patients are
included in a mortality and morbidity screening combining individual
review and specific audit filters by specialized nurses. The patients
who are identified as having higher probability for OFI are discussed in
a multidisciplinary conference held every 6-8 weeks. Examples of OFI
include the need for better organization or more senior members taking
part in the trauma care. The presence or absence of OFI is decided among
the participants of the conference and then recorded in the trauma care
quality database.

## Participants

The trauma registry includes all patients admitted with trauma team
activation, regardless of Injury Severity Score (ISS), as well as
patients admitted without trauma team activation but found to have an
ISS of more than 9. We included all patients who had been included in
the morbidity and mortality screening process between January 1, 2014
and February 1, 2023 and who were also admitted to the ICU. We excluded
patients who were younger than 15 years and patients who were dead on
arrival.

## Variables and data sources/measurements

### Study outcome

The study outcome is on the presence of OFI in adult patients treated in
the ICU based on patient and process factors.

### Patient and process factors

The variables were factors chosen from the trauma registry, based on the
locally used audit filters, standard epidemiological factors and factors
registered in the Swedish Trauma registry. The categorical factors were
sex, age, if the patient was intubated or not and how many days, Revised
Trauma Score (RTS), Injury Severity Score (ISS), on call, time from
arrival at the hospital until first computed tomography (CT), days in
the ICU, ASA preinjury and survival after 30 days. On duty was defined
as before 8.00 a.m., after 5 p.m., or during a weekend - defined as
Saturday or Sunday.

## Study size

There were 14,022 patients in the trauma registry between January 1 2013
and February 1 2023. 2679 of these patients were admitted to the ICU. 3
patients were then excluded as they were younger than 15 years and 6
patients were excluded because they were dead on arrival. 928 patients
were excluded because there was no registration on whether there was an
OFI or not. 293 patients were excluded because they had missing values
in one or more variables. This leaves us with 143 outcomes for OFI, and
1,306 outcomes for "No OFI".

## Statistical methods

Bivariable logistic regression was used to determine unadjusted
associations and multivariable logistic regression to determine adjusted
associations between patient and process factors and OFI. Odds ratios
(OR) with associated 95% confidence intervals was used with a 5%
significance level.

## Ethical considerations

There are several ethical aspects to take into consideration regarding
this study: voluntary participation, informed consent, anonymity,
confidentiality, potential for harm and results communication.

As described previously, all patients with trauma team activation or
high ISS scores are included in the database used for this study. Having
suffered severe trauma might not be something the patients or their
relatives would want to be public information as the collected data
contain for instance vital parameters and demise as well as type of
trauma and time of arrival to the hospital, which might be sensitive
information. Therefore, the ethical aspect of voluntary participation
and informed consent has to be considered. The data in national
databases is collected from all patients registered and is not based on
consent, even though it might be informed, which is a drawback.

To respect the anonymity and confidentiality aspect the statistical
analysis was first conducted on synthetic data and later implemented on
the data collected from the trauma registry and the trauma care quality
database in order to protect personal numbers and their associated
information. The anonymity might ease the discomfort among the involved
patients and their relatives.

There is no direct potential for harm by being included in the study as
no intervention is made and it is an observational study.

In regard to results communication, the results from research based on
the data as well as a compilation of the original data is public. The
data being accessible for research does not directly benefit the
affected individual; however, it contributes greatly to the development
of the hospital care in Sweden, making it more efficient and equal.
Studies performed on data from the Karolinska University Hospital in
Solna are made to improve local trauma care systems. This way, it can
also benefit the society as improved local trauma care decreases trauma
death rate and improves quality of care and thus patient safety.

Based on these arguments it can be concluded that the advantages
outweigh the disadvantages. The advantages being the contribution to the
development of trauma systems which long-term improves patient safety
and equality, and that the study protects the individual's anonymity and
confidentiality by using synthetic data and later implementing it on the
real data. The disadvantage being the potential emotional harm from the
absence of voluntary participation.

The study was approved by the Swedish Ethical Review Authority, approval
numbers 2023-02975-02, 2021-03531, 2021-02541.

# Results

## Participants

```{r, include=FALSE}
source("functions/flowchart.R")
```

```{r, echo=FALSE}
for (i in 1:(length(vert) - 1)) {
  connectGrob(vert[[i]], vert[[i + 1]], type = "vert") %>%
    print
}

connectGrob(vert$eligible, excluded, type = "L")
connectGrob(vert$included, excluded1, type = "L")

vert
excluded
excluded1
```

14,022 patients were included in the trauma registry and trauma care
quality database between 2017 and 2023. 12,278 patients were excluded,
leaving a total of 1,742 patients eligible for the study. 11,343 were
excluded because they were not treated in the ICU. 3 patients were
excluded because they were under 15 years old. 6 patients were excluded
because they were dead on arrival, and 928 patients were excluded
because there was no data on the presence or absence OFI. 293 patients
were excluded because there was missing data, leaving a total of 1,449
patients included in the study.

## Descriptive data

```{r include_scripts, include=FALSE}
source("functions/table1.R")
```

```{r, echo=FALSE}
table2
```

Table 1 present sample characteristics. 1,096 (76%) patients were male
and 353 (24%) patients were female. 846 (58%) patients had no
intubation, 241 (17%) patients were intubated for 1-7 days and 113
(7,8%) patients were intubated for more than 7 days. The number of
patients who arrived outside working hours --- on Saturday or Sunday or
after 5 pm or before 8 am were 1,042 (72%). 1,064 (73%) patients were
equal to or less than 7 days in the ICU, and 385 (27%) patients were
more than 7 days in the ICU. 1,080 (75%) patients had an ASA-score of
1-2 before arrival to the hospital, and 369 (25%) patients had an
ASA-score of 3-7 before arrival to the hospital. 286 (20%) patients died
within 30 days after hospitalization.

## Main results

```{r, include=FALSE}
source("functions/table2.R")
```

```{r, echo=FALSE}
table3_merge
```

Table 2 shows the unadjusted and adjusted associations of the selected
patient and process factors in relation to OFI. In the unadjusted
analysis the factors that were significantly associated with the
presence of OFI were: Revised Trauma Score
`r inline_text(table3a, variable = RTS)`, ≤ 7 days in the ICU
`r inline_text(table3a, variable = daysinICU, level = "≤ 7 days")` and
death within 30 days after hospitalization
`r inline_text(table3a, variable = Survival, level = "Dead")`.

In the adjusted analysis the following factors demonstrated significant
association with OFI: Revised Trauma Score
`r inline_text(table3b, variable = RTS)` and ≤ 7 days in the ICU
`r inline_text(table3b, variable = daysinICU, level = "≤ 7 days")`.

RTS was significantly associated with increased odds of OFI in both the
unadjusted and adjusted analyses. ≤ 7 days in the ICU was significantly
associated with decreasing odds of OFI in both the unadjusted and
adjusted analyses.

Sex, Age, Intubation, ISS, time to first CT, on duty and ASA preinjury
were not significantly associated with OFI in the unadjusted nor the
adjusted analyses. Female sex, intubation \> 7 days, \> 7 days in the
ICU, ASA preinjury 1-2 and survival 30 days after hospitalisation could
not be analysed with logistic regression.

```{r, include=FALSE}
source("functions/table3.R")
```

```{r, echo=FALSE}
table3alive_merge
```

Table 3 shows the unadjusted and adjusted associations of OFI and the
selected patient and process factors among the patients who were alive
30 days after hospitalization. In the unadjusted analysis time to first
CT `r inline_text(table3aalive, variable = TimeFCT)` was significant.
Indicating that it has no correlation with increasing or decreasing odds
of OFI.

# Discussion

The key results from this study are that we found that higher RTS was
associated with higher odds of OFI in both the adjusted and unadjusted
analyses, ≤ 7 days in the ICU was associated with decreased risk of OFI
in both the unadjusted and adjusted analyses, and death within 30 days
after hospitalization was associated with OFI in the unadjusted, but not
in the adjusted analysis. Age, gender, intubation, ISS, time to first
CT, on duty and ASA preinjury were not significantly associated with OFI
in neither the unadjusted or the adjusted analyses. In the subgroup
analysis of patients who were alive 30 days after hospitalization time
to first CT showed significance for having no relation to OFI.

We found that higher RTS was associated with higher odds of OFI. RTS is
a combined value of systolic blood pressure, Glasgow Coma Scale and
respiratory rate with more weight towards GCS. An explanation for the
increased odds of OFI in patients with increased RTS could be that the
patients with low RTS values - with low GCS, RR and SBP also are the
ones that are in a bad state upon arrival --- upon which emergency
procedures e.g. intubation have been carried out rapidly due to the
patients being visibly ill from the beginning. The patients who are
being missed and where certain procedures take longer time before being
performed are the patients who seem stable upon arrival, but who later
become worse --- the patients with comparatively higher RTS scores. In
Albaaj's study respiratory rate \>29, systolic blood pressure 50-75 mmHg
and 76-89 mmHg, and Glasgow Coma Scale 9-12 were significantly related
to increased odds of OFI in the unadjusted analysis, but not in the
adjusted analysis among patients from the part of the Swedish Trauma
Registry covering patients from the Karolinska University Hospital in
Solna [@12]. All of these variables equal medium to higher RTS values
--- capturing the group of patients who's vital parameters are
negatively affected by the trauma, but who are relatively stable in the
group of trauma patients. This creates a sense of false security,
leading to increased time until necessary emergency procedures are
performed.

The variable ≤ 7 days in the ICU was significantly associated with
decreasing odds of OFI in both the adjusted and unadjusted analyses.
There are two aspects to consider concerning this variable. One aspect
would be that less days spent in the ICU compared to more days spent in
the ICU indicates that patients have been in a better state and needed
less time of recovery before leaving the hospital. Furthermore, a longer
stay leads to more space for mistakes being made. The other aspect of
the variable would be that few days in the ICU could indicate the
patient passing away before any mistakes having been made due to the
patient being very ill from the beginning --- cases in which the deaths
were not preventable, but rather expected.

Previous studies have shown that prolonged ICU stay is a risk factor for
increased mortality and morbidity in trauma patients [@22, @26].
Something to consider is the reason for the lengthened stay in the ICU.
Studies have shown that patient characteristics on admission, like
comorbidities, age and severity of illness do not fully explain the
increased long-term mortality. With lengthened stay in the ICU the
reason for stay in the ICU become increasingly less associated with
patient's original reason for admittance and increasingly associated
with persistent chronic illness with associated organ dysfunction [@23,
@24]. This being the case, this might be an explanation for the lack of
sigificance for ASA preinjury and ISS in the analysis. As stated above
severity of illness and comorbidities do not fully explain increased
long-term mortality. OFI may therefore not correlate fully with ASA
preinjury and ISS respectively which are factors describing
comorbidities and the severity of illness.

However, a retrospective cohort has shown that lower admission illness
severity and younger age were associated with earlier discharge and
therefore less associated with persistent critical illness, which would
then indicate a stronger association between ISS and ASA preinjury to
length of stay in the ICU as the patients who are in a better state
leave the hospital earlier and there is less odds for OFI. On the other
hand, the cohort also showed that lower comorbidity was associated with
persistent critical illness rather than early discharge from the ICU ---
supporting the opposite side of the argument [@25].

Several studies have also studied the association between different
factors that affect the length of stay in the ICU. Associations between
hypoglycaemia, use of any corticosteroids, use of any neuromuscular
agents, use of any benzodiazepines for more than 1 day, the need for
mechanical ventilation for \> 2 days, initiation of new dialysis and the
occurrence of new infection has been found to have a correlation to
increased total and 5-year mortality [@23]. An association between
bleeding in patients and increased length of stay in the ICU has also
been found [@27].

Death within 30 days after hospitalization was associated with OFI in
the unadjusted analysis but not in the adjusted analysis. Patients who
died within 30 days after hospitalization could either have suffered a
severe trauma or have been in a poor state from the beginning; therefore
increasing the odds of OFI.

Sex and age were not significant in the unadjusted nor the adjusted
analyses, which implicates that the patients were not treated
differently based on age or gender. There are many studies on gender in
association to health care. It has previously been shown that there is
no difference in willingness to admit to the ICU between cases
describing a man or woman in the physician's responses [@29]. However,
studies also show that, among severely injured patients, female gender
is associated with lower in-hospital mortality among those aged 16-44
years. Female gender is also associated with decreased likelihood for
ICU admission [@30]. Male gender contributes more to poor outcomes in
postoperative ICU patients, and two-thirds of ICU resources are consumed
by male patients [@31]. However, in our study we could not see any
correlation between OFI and one specific gender in the occurrence of
OFI. Although we can also see that the majority of the patients included
in this study were also males. There is possibly a reason for higher
admittance of males to the ICU and association with poorer outcomes, but
it has nothing to do with odds of OFI.

ASA preinjury was not significant in the unadjusted nor the adjusted
analyses. Preinjury ASA score has been shown to be a reliable score for
predicting mortality after trauma, predicting readmission after
traumatic injury and for classifying comorbidity in trauma patients
[@32, @33, @34]. ASA score was originally developed to assess a
patient's health before surgical procedures and therefore provided
information on how much risk for postoperative complications there was.
Studies have investigated if ASA score also is a good predictor in other
aspects of medical care and as discussed above it is a predictor for
mortality after trauma and for readmission. However, this does not
automatically mean that there is an correlation between ASA score and
OFI. Patient's with higher ASA scores might receive more attention when
admitted to the ICU due to their comorbidities, but as the variable used
in this study is ASA preinjury --- a measurement of the patient's health
before the trauma --- this might not contribute to a fair picture of the
current state of the patient, which injury scores like RTS might.

The variable on duty was not significant in either analyses indicating
that the time of arrival of the patients did not affect their medical
treatment. Trying to reduce number of OFI can therefore not be solved by
increasing medical resources outside working hours. On-call sleeping has
been studied and it has been stated that sleeping between calls is
beneficial for waking performance. However, there might be a temporary
increase in performance impairment that should be taken into
consideration [@35].

Something to consider is why ISS was not significant in the analyses.
Patients admitted to the ICU are usually in a bad shape with generally
high ISS. ISS is based on the square of the three highest AIS scores
allocated to the three most severly injured body parts. The downside of
ISS on its own is that it has higher predictive value when combined with
other parameters like vital signs, comorbidities and mechanism of
injury. ISS is an anatomical model, compared to RTS which is a
physiological model. Among trauma patients admitted to the ICU the
physiological models perform better than the anatomical models. However,
patients with low scores in the physiological model also require an
anatomic analysis of the injuries to determine their severity [@36].
NISS has been found to be better than ISS in predicting the ICU
admission of trauma patients, as well as for predicting prolonged ICU
length of stay [@37]. This could be an explanation to why ISS is not
significant in the results of this analysis, as RTS as well as NISS are
potentially better models for patients admitted to the ICU.

### Strengths and limitations

One of the strengths of this study is that it is a registry-based study
which means that the quality of the data is reliable. The collection of
the included variables were pre-defined which means the reporting of the
included variables was standardized. Furthermore, the specific inclusion
criteria of the study: including patients 15 years and older, who were
not dead on arrival and who were admitted to the ICU, resulted in a
specific and well-defined cohort. This contributes to the reliability
and reproducibility of the study.

There were several limitations to our study: the limited study size, the
selection of patients for the morbidity and mortality review and the
definition and use of the physiological variables of the study.

Firstly, the study size was limited. This was a single-centre study and
illustrate the situation at the Karolinska Intitute Hospital in Solna.
Due to the small cohort of this study, it limits the applicability of
this study to other level 1 trauma centers and ICUs. The small cohort
might also affect the accuracy of the results.

Secondly, the selection of patients for the morbidity and mortality
review relied mostly on the local audit filters used at the Karolinska
Institute Hospital in Solna. Using audit filters to select patients for
the conference is associated with high false positive rates, ranging
from 24% to 80% [@38]. Even though the use of audit filters is
associated with high false positive rates, there is still a risk of
misplacement where some patients with OFI were not selected by for the
review based on the audit filters although there was an OFI that
occurred later on during the hospital stay.

The definition of the RTS-value is a combination of GCS, SBP and RR. The
variables used were the values from the emergency department. When these
were missing, the prehospital values were used. If these were also
missing, but there was documentation indicating that the patients had
been intubated, we replaced GCS and RR with a value of 0. This is
motivated for due to patients not breathing or being conscious when
intubated. The acquired values were then combined to calculate a value
for RTS. If the patient had missing values for either one of GCS, SBP or
RR they were excluded from the study. The handling of these variables
and their combination to RTS could affect the result as the values used
are from either prehospital values, emergency department values or from
identification of intubation. This could potentially affect the accuracy
of the data and decrease the possibility for reproducibility. However,
one could also argue that as the RTS value is a combined score of 3
physiological variables the combination of them decreases the overall
inaccuracy and produces a satisfactory measurement of the physiological
state of the patient.

### Generalisability

The study is conducted on 143 outcomes of OFI. There were few outcomes
because we included only those who were admitted to the ICU, \> 14 years
old, not dead on arrival and had a registration for the presence or
absence of OFI. The external validity is therefore compromised due to
the limited size of the study. The internal validity is however high
thanks to the well-defined cohort specifying trauma patients admitted to
the ICU at the Karolinska University Hospital in Solna. The results from
this study is therefore highly relevant for understanding where further
efforts for trauma improvement should be directed among ICU trauma
patients at this hospital.

### Implications

The potential clinical implication for this study is to further
contribute to the development of improved trauma care. Identifying the
group of patients admitted to the ICU is another step in the direction
of identifying the patient's who would benefit the most from further
improvements in trauma care. Trauma has a great global impact and this
study amongst others on the subject all contribute to the more specified
efforts in trauma quality improvement systems. Additionally, this study
has offered further insights to the patient profile of the trauma center
at Karolinska University Hospital in Solna admitted to the ICU.

### Future studies

More research is needed on the area of OFI in patients admitted to the
ICU. The most relevant limitation of this study was the limited study
size. A suggestion would be to conduct a multi-centre study, covering
larger databases or combining several databases, on this group of
patients to further investigate areas for improvement and to further
contribute to the improvement of trauma quality programs.

Based on the results from this study, variables of interest for future
studies could be number of days in the ICU and investigation of GCS, RR
and SBP separately instead of them as a combined RTS-score. It would be
interesting to see if one of the vital parameters is better for
predicting OFI than the others in a larger cohort.

Based on the statistical significance for higher RTS being associated
with higher odds of OFI, efforts should be more focused towards trauma
patients who are relatively stable upon arrival, but who have a risk of
getting worse. Future studies should investigate what factors that could
potentially indicate eventual worsening of the state of the patient.
Focusing on patients who are stable upon arrival overall is very
extensive, which further emphasizes the importance of identifying such
variables.

ASA preinjury did not show any correlation with OFI in this study.
However, it would be interesting to further investigate the correlation
between ASA preinjury or another scoring system of the patient's general
health before admittance to the hospital, and a combined physiological
and anatomical injury scoring system to see if they predict OFI better
combined than separately.

# Conclusion

The aim of this study was to characterize patient and process factors
among adult patients admitted to the ICU that were associated with OFI.
The study showed that higher RTS score is associated with increased odds
of OFI indicating that patients who are relatively well upon arrival to
the hospital are the group where improved trauma care should get more
focus than they do now. The logistical regression also showed that ≤ 7
days in the ICU was associated with decreased odds of OFI indicating
that patients who are at the hospital less days in the ICU compared to
more days in the ICU are less prone to opportunities for improvement.

# Contributions

This study has been conducted by Elin Sun Cao with continuous guidance
and feedback provided by Martin Gerdin Wärnberg, Jonatan Attergrim and
Kelvin Szolnoky.

# Acknowledgements

I want to extend my greatest gratitude to my supervisor Martin Gerdin
Wärnberg and co-supervisors Jonatan Attergrim and Kelvin Szolnoky who
have provided me with guidance and support throughout this semester.
Their extensive knowledge, deep insight into the subject and
availability and support has greatly contributed to the development of
this project. Additionally, I want to express my gratitude to my friends
and family who have been there contributing with encouragement and
positive energy.

# References

<!-- Do not edit by hand, references will be inserted and formatted automatically once you knit this document -->